---
title: "2021-07-26_ArchR-comparison"
author: "Jane Kim"
date: "7/26/2021"
output: html_document
---

```{r}
library(ArchR)
```


```{r}
addArchRThreads(threads = 4)
addArchRGenome("hg38")
```
Follow the instructions here: https://greenleaflab.github.io/ArchR_2020/Ex-Analyze-Multiome.html

Set Working Directory to: ~/Desktop/HARVARD/Dana_Farber_Research/comparison
Inside the working directory create a directory named pbmc_10k
In the directory pbmc_10k download the following three files:
```{bash}
wget https://cf.10xgenomics.com/samples/cell-arc/1.0.0/pbmc_granulocyte_sorted_10k/pbmc_granulocyte_sorted_10k_atac_fragments.tsv.gz
wget https://cf.10xgenomics.com/samples/cell-arc/1.0.0/pbmc_granulocyte_sorted_10k/pbmc_granulocyte_sorted_10k_atac_fragments.tsv.gz.tbi
wget https://cf.10xgenomics.com/samples/cell-arc/1.0.0/pbmc_granulocyte_sorted_10k/pbmc_granulocyte_sorted_10k_filtered_feature_bc_matrix.h5

```


Get input file
```{r}
#Get Input Fragment Files
inputFiles <- getInputFiles(paths = "/Users/jeongeunkim/Desktop/HARVARD/Dana_Farber_Research/comparison/pbmc_10k")[1]
names(inputFiles) <- "PBMC_10k"
```

```{r}
ArrowFiles <- createArrowFiles(
  inputFiles = inputFiles,
  sampleNames = names(inputFiles),
  filterTSS = 10, #Dont set this too high because you can always increase later
  filterFrags = 10^3.5, 
  addTileMat = TRUE,
  addGeneScoreMat = TRUE
)
```
```{r}
proj <- ArchRProject(
  ArrowFiles = ArrowFiles, 
  outputDirectory = "benchmark",
  copyArrows = TRUE #This is recommened so that if you modify the Arrow files you have an original copy for later usage.
)
```
```{r}
seRNA <- import10xFeatureMatrix(
    input = c("/Users/jeongeunkim/Desktop/HARVARD/Dana_Farber_Research/comparison/pbmc_10k/filtered_feature_bc_matrix.h5"),
    names = c("PBMC_10k")
)
```
```{r}
seRNA
```
```{r}
proj <- addGeneExpressionMatrix(input = proj, seRNA = seRNA, force = TRUE)
```
```{r}
projfilter1 <- proj[proj$TSSEnrichment > 6 & proj$nFrags > 2500 & !is.na(proj$Gex_nUMI)]
```

```{r}
projfilter1 <- addIterativeLSI(
    ArchRProj = projfilter1, 
    clusterParams = list(
      resolution = 0.2, # change this to 0.6 
      sampleCells = 10000,
      n.start = 10
    ),
    saveIterations = FALSE,
    useMatrix = "TileMatrix", 
    depthCol = "nFrags",
    name = "LSI_ATAC"
)
```

```{r}
projfilter1 <- addIterativeLSI(
    ArchRProj = projfilter1, 
    clusterParams = list(
      resolution = 0.2, 
      sampleCells = 10000,
      n.start = 10
    ),
    saveIterations = FALSE,
    useMatrix = "GeneExpressionMatrix", 
    depthCol = "Gex_nUMI",
    varFeatures = 2500,
    firstSelection = "variable",
    binarize = FALSE,
    name = "LSI_RNA"
)
```
```{r}
projfilter1 <- addCombinedDims(projfilter1, reducedDims = c("LSI_ATAC", "LSI_RNA"), name =  "LSI_Combined")
```

Run UMAPs
```{r}
projfilter1 <- addUMAP(projfilter1, reducedDims = "LSI_ATAC", name = "UMAP_ATAC", minDist = 0.8, force = TRUE) #default n.neighbors = 40
projfilter1 <- addUMAP(projfilter1, reducedDims = "LSI_RNA", name = "UMAP_RNA", minDist = 0.8, force = TRUE)
projfilter1 <- addUMAP(projfilter1, reducedDims = "LSI_Combined", name = "UMAP_Combined", minDist = 0.8, force = TRUE)
```
Add clusters
```{r}
projfilter1 <- addClusters(projfilter1, reducedDims = "LSI_Combined", name = "Clusters", resolution = 0.4, force = TRUE)
```
Plot Embeddings
```{r}
p1 <- plotEmbedding(projfilter1, name = "Clusters", embedding = "UMAP_ATAC", size = 1.5, labelAsFactors=F, labelMeans=F)
p2 <- plotEmbedding(projfilter1, name = "Clusters", embedding = "UMAP_RNA", size = 1.5, labelAsFactors=F, labelMeans=F)
p3 <- plotEmbedding(projfilter1, name = "Clusters", embedding = "UMAP_Combined", size = 1.5, labelAsFactors=F, labelMeans=F)
```

```{r}
p1+p2+p3
```
```{r}
p1
```
```{r}
p2
```
```{r}
p3
```
Add 2 more clusters: using reducedDims = "LSI_ATAC" and "LSI_RNA"
```{r}
projfilter1 <- addClusters(projfilter1, reducedDims = "LSI_ATAC", name = "Clusters_ATAC", resolution = 0.4, force = TRUE)
projfilter1 <- addClusters(projfilter1, reducedDims = "LSI_RNA", name = "Clusters_RNA", resolution = 0.4, force = TRUE)
```

Visualization with Clusters_RNA
```{r}
p4 <- plotEmbedding(projfilter1, name = "Clusters_RNA", embedding = "UMAP_ATAC", size = 1.5, labelAsFactors=F, labelMeans=F)
p5 <- plotEmbedding(projfilter1, name = "Clusters_RNA", embedding = "UMAP_RNA", size = 1.5, labelAsFactors=F, labelMeans=F)
p6 <- plotEmbedding(projfilter1, name = "Clusters_RNA", embedding = "UMAP_Combined", size = 1.5, labelAsFactors=F, labelMeans=F)
```
```{r}
p4+p5+p6
```
Visualization with Clusters_ATAC
```{r}
p7 <- plotEmbedding(projfilter1, name = "Clusters_ATAC", embedding = "UMAP_ATAC", size = 1.5, labelAsFactors=F, labelMeans=F)
p8 <- plotEmbedding(projfilter1, name = "Clusters_ATAC", embedding = "UMAP_RNA", size = 1.5, labelAsFactors=F, labelMeans=F)
p9 <- plotEmbedding(projfilter1, name = "Clusters_ATAC", embedding = "UMAP_Combined", size = 1.5, labelAsFactors=F, labelMeans=F)
```

```{r}
p7+p8+p9
```
```{r}
saveArchRProject(ArchRProj = projfilter1, outputDirectory = "Save-Projfilter1", load = TRUE)
```

Re-running with resolution = 0.6
```{r}
projfilter1 <- addIterativeLSI(
    ArchRProj = projfilter1, 
    clusterParams = list(
      resolution = 0.6, # change this to 0.6 
      sampleCells = 10000,
      n.start = 10
    ),
    saveIterations = FALSE,
    useMatrix = "TileMatrix", 
    depthCol = "nFrags",
    name = "LSI_ATAC_res0.6"
)
```

```{r}
projfilter1 <- addIterativeLSI(
    ArchRProj = projfilter1, 
    clusterParams = list(
      resolution = 0.6, 
      sampleCells = 10000,
      n.start = 10
    ),
    saveIterations = FALSE,
    useMatrix = "GeneExpressionMatrix", 
    depthCol = "Gex_nUMI",
    varFeatures = 2500,
    firstSelection = "variable",
    binarize = FALSE,
    name = "LSI_RNA_res0.6"
)
```

```{r}
projfilter1 <- addCombinedDims(projfilter1, reducedDims = c("LSI_ATAC_res0.6", "LSI_RNA_res0.6"), name =  "LSI_Combined_res0.6")
```

Run UMAPs
```{r}
projfilter1 <- addUMAP(projfilter1, reducedDims = "LSI_ATAC_res0.6", name = "UMAP_ATAC_res0.6", minDist = 0.8, force = TRUE)
projfilter1 <- addUMAP(projfilter1, reducedDims = "LSI_RNA_res0.6", name = "UMAP_RNA_res0.6", minDist = 0.8, force = TRUE)
projfilter1 <- addUMAP(projfilter1, reducedDims = "LSI_Combined_res0.6", name = "UMAP_Combined_res.0.6", minDist = 0.8, force = TRUE)
```

```{r}
projfilter1 <- addClusters(projfilter1, reducedDims = "LSI_Combined_res0.6", name = "Clusters_res0.6", resolution = 0.6, force = TRUE)
```

```{r}
p10 <- plotEmbedding(projfilter1, name = "Clusters_res0.6", embedding = "UMAP_ATAC_res0.6", size = 1.5, labelAsFactors=F, labelMeans=F)
p11 <- plotEmbedding(projfilter1, name = "Clusters_res0.6", embedding = "UMAP_RNA_res0.6", size = 1.5, labelAsFactors=F, labelMeans=F)
p12 <- plotEmbedding(projfilter1, name = "Clusters_res0.6", embedding = "UMAP_Combined_res.0.6", size = 1.5, labelAsFactors=F, labelMeans=F)
```
```{r}
p10+p11+p12
```

```{r}
saveArchRProject(ArchRProj = projfilter1, outputDirectory = "Save-Projfilter2", load = TRUE)
```

```{r}
projfilter1 <- addClusters(projfilter1, reducedDims = "LSI_ATAC_res0.6", name = "Clusters_binmethod_res0.6", resolution = 0.6, force = TRUE)
```

```{r}
p13 <- plotEmbedding(projfilter1, name = "Clusters_binmethod_res0.6", embedding = "UMAP_ATAC_res0.6", size = 1.5, labelAsFactors=F, labelMeans=F)
p14 <- plotEmbedding(projfilter1, name = "Clusters_binmethod_res0.6", embedding = "UMAP_RNA_res0.6", size = 1.5, labelAsFactors=F, labelMeans=F)
p15 <- plotEmbedding(projfilter1, name = "Clusters_binmethod_res0.6", embedding = "UMAP_Combined_res.0.6", size = 1.5, labelAsFactors=F, labelMeans=F)
```
```{r}
p13
```
```{r}
projfilter1 <- addClusters(projfilter1, reducedDims = "LSI_Combined", name = "test", resolution = 0.6, force = TRUE)
```
```{r}
projfilter1 <- addClusters(projfilter1, reducedDims = "LSI_ATAC_res0.6", name = "Clusters_bin_method_0.6", resolution = 0.6, force = TRUE)
```

```{r}
tileMatrixtest <- getMatrixFromProject(
  ArchRProj = projfilter1,
  useMatrix = "TileMatrix",
  binarize = TRUE)
```

